---
title: "Supplemental Materials"
shorttitle: "Smiling and happiness: Supplement"
author:
  - name: Nicholas A. Coles
    corresponding: true
    email: ncoles@ufl.edu
    affiliations:
      - id: ufl
        name: University of Florida
        department: Department of Psychology
        address: 945 Center Dr
        city: Gainesville
        region: FL
        postal-code: "32611"
  - name: Annabel Dang
    affiliations:
      - ref: ufl
  - name: Joao Francisco Goes Braga Takayanagi
    affiliations:
      - ref: ufl
suppress-abstract: true
suppress-keywords: true
bibliography: references.bib
floatsintext: true
numbered-lines: false
mask: false
draft-date: false
meta-analysis: false
format:
  apaquarto-docx: default
  apaquarto-pdf:
    documentmode: doc
  apaquarto-html:
    knitr:
      opts_chunk:
        dev: png
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(knitr)

# Read processed data and the pairwise-comparison object exported by the main analysis script
df <- read_csv("../../../../data/smile25b_processed_data.csv", show_col_types = FALSE)
happy_simple <- readRDS("../../../../data/main_analysis/smile25b_happy_results_simple.Rds")

# Reporting helpers (mirrors smile_R3.qmd's setup chunk)
num <- function(x, digits = 2) formatC(x, format = "f", digits = digits)

fmt_bf <- function(bf) {
  ifelse(bf >= 1000, formatC(bf, format = "e", digits = 2), num(bf))
}

p_stars <- function(p) {
  case_when(
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ ""
  )
}

# Cell sizes: context x threat x repetition is a between-subjects grouping,
# so the same N applies to both the smile and natural pose scores within a cell
n_cell <- df %>%
  count(context, threat, repetition, name = "N")

si_table1 <- happy_simple %>%
  left_join(n_cell, by = c("context", "threat", "repetition")) %>%
  mutate(
    ci_low  = estimate - qt(.975, df) * SE,
    ci_high = estimate + qt(.975, df) * SE,
    `Emotional Context`     = context,
    `Movement Repetitions`  = repetition,
    Punishment              = if_else(threat == "Threat", "yes", "no"),
    `95% CI`                = paste0("[", num(ci_low), ", ", num(ci_high), "]"),
    EMM                     = num(estimate),
    p                       = p_stars(p.value),
    `BF~10~`                = fmt_bf(BF10),
    .context_order    = factor(context, levels = c("positive", "negative")),
    .repetition_order = factor(repetition, levels = c("one", "ten")),
    .threat_order      = factor(threat, levels = c("No threat", "Threat"))
  ) %>%
  arrange(.context_order, .repetition_order, .threat_order) %>%
  select(`Emotional Context`, `Movement Repetitions`, Punishment, N, EMM, `95% CI`, p, `BF~10~`)
```

# Supplemental Table 1

**Table S1**

*Pairwise Comparisons of Posed Smiles vs. Natural Expressions on Self-Reported Happiness*

```{r}
#| label: tbl-si1-output
#| echo: false

kable(
  si_table1,
  align = c("l", "l", "l", "r", "r", "c", "c", "r")
)
```

*Note.* EMM = estimated marginal mean of the pose difference (smile − natural), derived
from `emmeans` pairwise comparisons. CI = confidence interval. BF~10~ = Bayes factor for
the alternative hypothesis (a pose difference) over the null. \* *p* < .05. \*\* *p* <
.01. \*\*\* *p* < .001.

# Supplemental Figure 1

![Fig S1. Hard sensitivity check. Differences in self-reported happiness after participants posed happy vs. neutral expressions (y-axis), restricted to participants who complied with facial expression instructions, made no math errors, and were not aware of the study's true purpose. Points represent the estimated marginal mean and error bars represent 95% confidence intervals. Results are visualized separately for each combination of the three experimentally manipulated factors (denoted by x-axis and legend).](../../../../figures/smile25b_happy_sens_plot.jpg){#fig-si1}

# Supplemental Figure 2

![Fig S2. Soft sensitivity check. Differences in self-reported happiness after participants posed happy vs. neutral expressions (y-axis), restricted to participants who made no math errors, were not aware of the study's true purpose, and showed at least a modest increase (AU12 activation difference ≥ 0.5) in smiling during the posed vs. natural condition. Points represent the estimated marginal mean and error bars represent 95% confidence intervals. Results are visualized separately for each combination of the three experimentally manipulated factors (denoted by x-axis and legend).](../../../../figures/smile25b_happy_sens_soft_plot.jpg){#fig-si2}
