---
title: "Supplemental Materials: Smiling your way to happiness or misery? Experimental tests of competing perspectives"
shorttitle: "Supplement: Smiling and happiness"
author:
  - name: Nicholas A. Coles
    corresponding: true
    email: ncoles@ufl.edu
    affiliations:
      - id: ufl
        name: University of Florida
        department: Department of Psychology
        address: 945 Center Dr
        city: Gainesville
        region: FL
        postal-code: "32611"
  - name: Annabel Dang
    affiliations:
      - ref: ufl
  - name: Joao Francisco Goes Braga Takayanagi
    affiliations:
      - ref: ufl
suppress-abstract: true
suppress-keywords: true
bibliography: references.bib
floatsintext: true
numbered-lines: false
mask: false
draft-date: false
meta-analysis: false
format:
  apaquarto-docx: default
  apaquarto-pdf:
    documentmode: doc
  apaquarto-html:
    knitr:
      opts_chunk:
        dev: png
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(knitr)

# Read processed data and the pairwise-comparison object exported by the main analysis script
df <- read_csv("../data/smile25b_processed_data.csv", show_col_types = FALSE)
happy_simple <- readRDS("../data/main_analysis/smile25b_happy_results_simple.Rds")

# Reporting helpers (mirrors smile_R3.qmd's setup chunk)
num <- function(x, digits = 2) formatC(x, format = "f", digits = digits)

fmt_bf <- function(bf) {
  ifelse(bf > 100, "> 100", num(bf))
}

p_stars <- function(p) {
  case_when(
    p < .001 ~ "***",
    p < .01  ~ "**",
    p < .05  ~ "*",
    TRUE     ~ ""
  )
}

# Cell sizes: context x threat x repetition is a between-subjects grouping,
# so the same N applies to both the smile and natural pose scores within a cell
n_cell <- df %>%
  count(context, threat, repetition, name = "N")

# AU12 compliance exclusion counts (facial expression check only, independent
# of the math-error / awareness criteria applied in the main sensitivity analysis)
au12_compliance_hard_n   <- sum(!df$face_compliance_scalar, na.rm = TRUE)
au12_compliance_hard_pct <- au12_compliance_hard_n / nrow(df) * 100

au12_compliance_soft_n <- df %>%
  filter(
    (AU12_scalar_smile - AU12_scalar_natura) < 0.5 |
      face_detection_natura < 0.8 |
      face_detection_smile < 0.8
  ) %>%
  nrow()
au12_compliance_soft_pct <- au12_compliance_soft_n / nrow(df) * 100

si_table1 <- happy_simple %>%
  left_join(n_cell, by = c("context", "threat", "repetition")) %>%
  mutate(
    ci_low  = estimate - qt(.975, df) * SE,
    ci_high = estimate + qt(.975, df) * SE,
    `Emotional Context`     = context,
    `Movement Repetitions`  = repetition,
    Punishment              = if_else(threat == "Threat", "yes", "no"),
    `95% CI`                = paste0("[", num(ci_low), ", ", num(ci_high), "]"),
    EMM                     = num(estimate),
    p                       = p_stars(p.value),
    `BF~10~`                = fmt_bf(BF10),
    .context_order    = factor(context, levels = c("positive", "negative")),
    .repetition_order = factor(repetition, levels = c("one", "ten")),
    .threat_order      = factor(threat, levels = c("No threat", "Threat"))
  ) %>%
  arrange(.context_order, .repetition_order, .threat_order) %>%
  select(`Emotional Context`, `Movement Repetitions`, Punishment, N, EMM, `95% CI`, p, `BF~10~`)
```

In addition to the overview in the main manuscript, Table SI contains pairwise comparisons of self-reported happiness after participants posed smiles vs. expressed themselves naturally. 

**Table S1**

*Pairwise Comparisons of Posed Smiles vs. Natural Expressions on Self-Reported Happiness*

```{r}
#| label: tbl-si1-output
#| echo: false

kable(
  si_table1,
  align = c("l", "l", "l", "r", "r", "c", "c", "r")
)
```

*Note.* EMM = estimated marginal mean of the pose difference (smile − natural), derived
from `emmeans` pairwise comparisons. CI = confidence interval. BF~10~ = Bayes factor for
the alternative hypothesis (a pose difference) over the null. \* *p* < .05. \*\* *p* <
.01. \*\*\* *p* < .001.

Supplemental Figure 1 provides an overview of our primary results when participants were excluded for not exhibiting moderate increases in smiling activity during the happy posing trials. Using Facial Action Coding System classifications, this was operationalized as Action Unit 12 activity, which is associated with pulling the corner of the lips back towards the ear. For facial expression analysis, we originally planned to use Hume AI. However, we later switched to OpenFace 2.2.0 because Hume AI deprecated their expression analysis application before we completed our study. Because Hume AI (0-1) and OpenFace (0-5) have different scaling, we adjusted our initial .3/1 cutoff to 1.5/5. This led to an unexpectedly large number of exclusions (n = `r au12_compliance_hard_n`; `r num(au12_compliance_hard_pct, 0)`% of total sample). We thus conducted an additional sensitivity analysis with a less stringent cutoff of 0.5/5 (n = `r au12_compliance_soft_n` excluded; `r num(au12_compliance_soft_pct, 0)`% of total sample).

As shown in Supplemental Figure 1, Panel A, our main results replicated when using a less stringent cutoff. When using a more stringent cutoff, patterns were numerically similar (Supplemental Figure 1, Panel B). However, the mood-boosting effects of voluntarily posing smiles in negative contexts were no longer statistically significant.

# Supplemental Figure 1

![Fig S1. Results from sensitivity analysis wherein participants with weak smiles were included (Panel A) and excluded (Panel B). Y-axis denotes differences in self-reported happiness after participants posed happy vs. neutral expressions. Points represent the estimated marginal mean and error bars represent 95% confidence intervals. Results are visualized separately for each combination of the three experimentally manipulated factors (denoted by x-axis and legend).](../figures/smile25b_sens_plot.jpg){#fig-si1}
